Weak Lensing Shear Inference:A Hierarchy Bayesian Approach

### 2013-02-21 write-up.tex
update write-up

\date{02/16/2013}
\maketitle
\begin{abstract}
We present a Hierarchy Bayesian method for shear inference in Weak
Lensing survey. With this method we are able to infer synchronously
cosmic shear at different sky patches as well as the intrinsic shape
distribution. We show that this Hierarchy Bayesian method is more
accurate than the traditional average method and (prove to a large
extent )it is an unbiased estimation of shear.

@@

$|\frac{\partial(|\epsilon_{0}|,\phi_{0})}{\partial(|\epsilon|,\phi)}|$
is the Jacobi Determinant term. 
\begin{equation}
|\frac{\partial(|\epsilon_{0}|,\phi_{0})}{\partial(|\epsilon_{\ell}|,\phi_{\ell})}|=|\frac{\partial(|\epsilon_{0}|,\phi_{0})}{\partial(x_{0},y_{0})}*\frac{\partial(x_{0},y_{0})}{\partial(x_{\ell},y_{\ell})}*\frac{\partial(x_{\ell},y_{\ell})}{\partial(|\epsilon_{\ell}|,\phi_{\ell})}|=\frac{|\epsilon_{\ell}|}{|\epsilon_{0}|}*|\frac{\partial(x_{0},y_{0})}{\partial(x_{\ell},y_{\ell})}|
\end{equation}

$(x_{0},y_{0})$ and $(x_{\ell},y_{\ell})$ are the Cartesian coordinates
of $\epsilon_{0}$ and $\epsilon_{\ell}$ on the complex plane.

In this paper 
\begin{equation}
P(\epsilon_{0}|\alpha)=|\epsilon_{0}|^{p-1}(1-|\epsilon_{0}|)^{q-1}/Beta(p,q)
\end{equation}

We can see that when p<2 $L(g,\alpha|\epsilon_{\ell})$ will have
singular points at $g=\epsilon_{\ell}$.In this paper to avoid this
issue we restrict p always larger than 2.

@@

@@

The prior distribution we used on $\alpha$ is : 
\begin{equation}
Prior(\alpha)=exp(-\xi*\sum(\alpha_{i}-\alpha_{i-1})^{2})
\end{equation}

@@

Figure 10 shows how the mean square error of the Hierarchy Bayesian
method varies with the number of sky patches used,in comparison with
traditional average method and likelihood analysis with correct P($\epsilon$)
hard-set.from this figure we can see that the mean square error of
the Hierarchy Bayesian method decreases as the number of patches goes
up,while this trend is less significant as number of patches becomes
larger.Moreover,down to 8 patches this Hierarchy Bayesian method still
out perform the traditional average method. Meanwhile the likelihood
analysis with correct P($\epsilon$) shows higher accuracy than Hierarchy
Bayesian method. That is consistent with our speculation. When the
number of patches goes to infinity ,we will have correct estimation
on P($\epsilon$) and Hierarchy Bayesian method converges to likelihood
analysis with correct $P(\epsilon)$ hard-set,ignoring the difference
between our step function model and real beta function,of course.

@@

Figure 10:Mean square error VS Number of patches.In the upper graph
number of galaxies per patch is 8 while in the lowwer it's 16.The
Green line is mean square error of traditional average while the red
line is that of likelihood analysis with correct $P(\epsilon)$ .

@@

\section{Systematic Bias}

In this section,we measured the levels of possible systemmatic bias
led by this HB method.

@@

The accuracy in Weak Lensing measurement depends on both galaxies
shape measurement and shear inference period.In this paper our main
concern is how to infere shear in a HB method instead off traditional
averaging method and it does not involve the shape measurement part
.Therefore we assume we have perfect galaxies ellipticity measurement,and
all the possible inaccuracy is due to this HB shear inference method.

Using 10,000 data sets we are able to get the value of m and c for
both beta function and step function model.Each data set we use in
this section has 64 patches with 8 galaxies in eah patch.

The region for $g_{i}^{true}$ is -0.1\textasciitilde{}0.1 .

For beta function model :

\begin{tabular}{|c|c|c|}
\hline 
 & Expect Value  & STD\tabularnewline
\hline 
\hline 
m1  & -0.00420  & 0.00304\tabularnewline
\hline 
c1  & 1.66$*10^{-4}$  & 1.73$*10^{-4}$\tabularnewline
\hline 
m2  & -0.00386  & 0.00290\tabularnewline
\hline 
c2  & 2.30$*10^{-4}$  & 1.81$*10^{-4}$\tabularnewline
\hline 
\end{tabular}

For step function model:

\begin{tabular}{|c|c|c|}
\hline 
 & Expect Value  & STD\tabularnewline
\hline 
\hline 
m1  & 0.00502  & 0.00287\tabularnewline
\hline 
c1  & 2.05$*10^{-4}$  & 1.73$*10^{-4}$\tabularnewline
\hline 
m2  & 0.00156  & 0.00292\tabularnewline
\hline 
c2  & 2.14$*10^{-4}$  & 1.87{*}$10^{-4}$\tabularnewline
\hline 
\end{tabular}

\includegraphics[scale=0.5]{step_g1}

\includegraphics[scale=0.5]{g2_step} 
\begin{quote}
Figure 9: Linear fit of $\delta g_{i}-g_{i}$ relationship for step
function model. 
\end{quote}
As we can see from the data and graph no significant systematic bias
can be found using 10,000 data sets.With that amount of data we able
to narrow down m to $\sim10^{-3}$ and c to $\sim10^{-4}$.

In Table 1 we show the requirements on m, c for current, upcoming
and future cosmic shear surveys (Joe Zuntz ,axrive 2013)

>>>>

\date{02/19/2013}
\maketitle
\begin{abstract}
We present a Hierarchy Bayesian method for shear inference in Weak
Lensing survey. With this method we are able to infer synchronously
cosmic shear at different sky patches as well as the intrinsic shape
distribution. We show that this Hierarchy Bayesian method is more
accurate than the traditional average method and (prove to a large
extent )it is an unbiased estimation of shear.

@@

$|\frac{\partial(|\epsilon_{0}|,\phi_{0})}{\partial(|\epsilon|,\phi)}|$
is the Jacobean Determinant term. 
\begin{equation}
|\frac{\partial(|\epsilon_{0}|,\phi_{0})}{\partial(|\epsilon_{\ell}|,\phi_{\ell})}|=|\frac{\partial(|\epsilon_{0}|,\phi_{0})}{\partial(x_{0},y_{0})}*\frac{\partial(x_{0},y_{0})}{\partial(x_{\ell},y_{\ell})}*\frac{\partial(x_{\ell},y_{\ell})}{\partial(|\epsilon_{\ell}|,\phi_{\ell})}|=\frac{|\epsilon_{\ell}|}{|\epsilon_{0}|}*|\frac{\partial(x_{0},y_{0})}{\partial(x_{\ell},y_{\ell})}|
\end{equation}

$(x_{0},y_{0})$ and $(x_{\ell},y_{\ell})$ are the Cartesian coordinates
of $\epsilon_{0}$ and $\epsilon_{\ell}$ on the complex plane.This
Jacobean term is numerically computaed.

@@

In this section 
\begin{equation}
P(\epsilon_{0}|\alpha)=|\epsilon_{0}|^{p-1}(1-|\epsilon_{0}|)^{q-1}/Beta(p,q)
\end{equation}

We can see that when p<2 $L(g,\alpha|\epsilon_{\ell})$ will have
singular points at $g=\epsilon_{\ell}$.In this section to avoid this
issue we restrict p always larger than 2.

@@

The prior distribution we used on $\alpha$ is : 
\begin{equation}
Prior(\vec{\alpha}|\xi)=exp(-\xi\sum(\alpha_{i}-\alpha_{i-1})^{2})
\end{equation}

@@

Figur 10 shows the distribution of $g_{estimation}-g_{true}$.From
the figure we can see that it's a gaussian like distribution centrered
at 0.

\includegraphics[scale=0.5]{dg2}

Figure 11 shows how the mean square error of the Hierarchy Bayesian
method varies with the number of sky patches used,in comparison with
traditional average method and likelihood analysis with correct P($\epsilon$)
hard-set.from this figure we can see that the mean square error of
the Hierarchy Bayesian method decreases as the number of patches goes
up,while this trend is less significant as number of patches becomes
larger.Moreover,down to 8 patches this Hierarchy Bayesian method still
out perform the traditional average method. Meanwhile the likelihood
analysis with correct P($\epsilon$) shows higher accuracy than Hierarchy
Bayesian method. That is consistent with our speculation. When the
number of patches goes to infinity ,we will have correct estimation
on P($\epsilon$) and Hierarchy Bayesian method converges to likelihood
analysis with correct $P(\epsilon)$ hard-set,ignoring the difference
between our step function model and real beta function,of course.

@@

Figure 11:Mean square error VS Number of patches.In the upper graph
number of galaxies per patch is 8 while in the lowwer it's 16.The
Green line is mean square error of traditional average while the red
line is that of likelihood analysis with correct $P(\epsilon)$ .

One limitation of the step function model is that it needs to be \textit{regularized};
this is achieved by assigning a prior PDF for the step heights $\vec{\alpha}$.
We use the form $Prior(\vec{\alpha}|\xi)=exp(-\xi\sum(\alpha_{i}-\alpha_{i-1})^{2})$
when inferring the P($\epsilon$) and $g$. The choice of this functional
form corresponds to the assumption that the true P($\epsilon$) is
smooth, which seems fairly natural. This function has a hyperparameter,
$\xi$, that needs to be set somehow. If $\xi$ is too small, the
inferred P($\epsilon$) will be too noisy, and the accuracy of the
$g$ inference will decrease. Likewise, choosing an $\xi$ that is
too large, the inferred P($\epsilon$) will be too smooth, and the
data will be under-fitted. In principle one could infer $\xi$ form
the data at hand, but in practice we find that .

a flat prior PDF for $\xi$ will make the inferred $P(\epsilon)$
too noisy.Therefore to make the inferred $P(\epsilon)$ smooth enough
we still need to assign some prior PDF on $\xi$ that pushes $\xi$
to a larger region,and this prior PDF also need to be tuned. We find
that $\xi$ is proportional to the mean data number density and can
be approximated by the following empirical formula: 
\begin{equation}
\xi=N<\rho(|\epsilon|)>/256*\frac{Nbin}{10}
\end{equation}
Here $<\rho|\epsilon|>=\int\rho(|\epsilon|)^{2}d(|\epsilon|)$ corresponds
to the average of data number density weighted with number density,and
Nbin is the number of bins used in step function model. With the help
of this empirical formula, we can tune this prefactor $\xi$ by doing
some simulation which has similar data distribution with the real
data.

@@

\section{Systematical Bias}

$\ \ \ \ $In this section,we measured the levels of possible systemmatical
bias led by this HB method.

@@

The accuracy of Weak Lensing survey depends on both galaxy shape measurement
and shear inference .In this paper our main concern is how to infere
shear in a HB method instead off traditional averaging method and
it does not involve the shape measurement part .Therefore we assume
we have perfect galaxies ellipticity measurement,and all the possible
inaccuracy is due to this HB shear inference method.

We calculated the average of $g^{estimated}-g^{true}$ of 10,000 data
sets as described above.Each data set has 64 patches and 8{*}64 galaxies,and
the true P($\epsilon$) is Beta(2.8,2.8) .For both Beta function model
and step funtion model we then determin the best fitted parameters
m and c of fomula (15).Figure 9 shows an example of step funtion model
with prior prefator $\xi=3$.The results of the systematical bias
analysis are tabulated in table 1.

\includegraphics[scale=0.5]{g1bias}

\includegraphics[scale=0.5]{g2bias}
\begin{quote}
Figure 9: Linear fit of $\delta g_{i}-g_{i}$ relationship for step
function model.

\begin{tabular}{|c|c|c|c|}
\hline 
Beta Model & Expect Value  & $\sigma$ & S/N\tabularnewline
\hline 
\hline 
m1  & -0.00420  & 0.00304 & 1.38\tabularnewline
\hline 
c1  & 1.66$*10^{-4}$  & 1.73$*10^{-4}$ & 0.96\tabularnewline
\hline 
m2  & -0.00386  & 0.00290 & 1.31\tabularnewline
\hline 
c2  & 2.30$*10^{-4}$  & 1.81$*10^{-4}$ & 1.27\tabularnewline
\hline 
\end{tabular}

\begin{tabular}{|c|c|c|c|}
\hline 
Step Model & Expect Value  & $\sigma$ & S/N\tabularnewline
\hline 
\hline 
m1  & -6.09{*}$10^{-4}$ & 0.00239 & 0.25\tabularnewline
\hline 
c1  & 1.82$*10^{-4}$  & 1.44$*10^{-4}$ & 1.26\tabularnewline
\hline 
m2  & 0.00178  & 0.00281 & 0.63\tabularnewline
\hline 
c2  & 2.978$*10^{-4}$  & 1.71{*}$10^{-4}$ & 1.74\tabularnewline
\hline 
\end{tabular}

Table 1:Tabulated measures of multiplicative error <$m_{i}>$ and
additive error $<c_{i}>$and their standard deviations ,along with
their S/N ratio. 
\end{quote}
In Table 1 we show the requirements on m, c for current, upcoming
and future cosmic shear surveys (Joe Zuntz ,axrive 2013)
